### 2021-03-24 analysis-word-cloud.Rmd

---
title: "Word Cloud"
author: "Emi Tanaka"
date: "24/03/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(glue)
library(pluralize)
library(lubridate)
library(plotly)
library(ggwordcloud)
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache.path = "cache/")
```

```{r fns}
singularize2 <- function(x) {
    res <- singularize(x)
    res <- ifelse(res=="bia", "bias", res)
    ifelse(res=="baye", "bayes", res)
}
stop_words_rex <- c(paste0("^", stop_words$word, "$"), "^doi", "^[0-9.]+$")
get_ngram <- function(ngram, desc = c("Title", "Description")) {
  desc <- match.arg(desc)
  words <- paste0("word", 1:ngram)
  doe_db %>% 
    unnest_tokens(word, desc, token = "ngrams", n = ngram) %>% 
    separate(word, words, sep = " ") %>% 
    mutate(across(num_range("word", 1:ngram), singularize2)) %>% 
    distinct(!!!map(c("Package", words), as.name)) %>% 
    # allow stop words in between
    filter(if_all(num_range("word", c(1, ngram)), 
                  ~!str_detect(., paste0(stop_words_rex, collapse = "|")))) %>% 
    count(!!!map(words, as.name), sort = TRUE) %>% 
    mutate(word = do.call("paste", map(words, ~eval(parse(text = .x))))) %>% 
    select(word, n) 
}

combine_dl <- function(ngram_df, desc) {
  dldat_sum %>% 
    left_join(doe_db, by = c("package" = "Package")) %>% 
    select(package, total, {{desc}}) %>% 
    mutate(bigram_title = map({{desc}}, ~filter(ngram_df, 
                                              n > 1,
                                              str_detect(tolower(.x), word)))) %>% 
    unnest_longer(bigram_title) %>% 
    unpack(bigram_title) %>% 
    group_by(word) %>% 
    summarise(n = unique(n),
              pkgs = list(package),
              total = sum(total)) %>% 
    filter(!is.na(word)) %>% 
    arrange(desc(total))
}
```

```{r ngrams, }
bigram_title <- get_ngram(2, "Title")
bigram_desc <- get_ngram(2, "Description")
trigram_desc <- get_ngram(3, "Description")
```

```{r wordcloud, fig.height = 3.3}
dldat_comb_bigram_title <- combine_dl(bigram_title, Title)
dldat_comb_bigram_desc <- combine_dl(bigram_desc, Title)
dldat_comb_trigram_desc <- combine_dl(trigram_desc, Title)

set.seed(1)
ggplot(rbind(dldat_comb_bigram_desc,
             dldat_comb_trigram_desc),
       aes(label = word, size = total, color = n)) +
  geom_text_wordcloud(shape = "square",
                      show.legend = TRUE) +
  scale_size_area(max_size = 8) +
  guides(size = FALSE) +
  scale_color_continuous(breaks = 1:11) +
  theme(legend.position = "bottom")
```