### 2021-03-30 analysis-pkg-selection.Rmd
selection of rpkgs

---
title: "Selection of R-packages for review"
author: "Emi Tanaka"
date: "24/03/2021"
output: html_document
---

```{r library, include=FALSE}
library(tidyverse)
library(magrittr)
library(glue)
library(kableExtra)
library(lubridate)
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache.path = "cache/")
```

```{r cache, eval = FALSE}
fs::dir_ls("cache") %>% 
  str_subset(".RData$") %>% 
  walk(load)
```

```{r ctv}
doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
                                              repos = "http://cran.rstudio.com/")[[1]]
survey_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
                                              repos = "http://cran.rstudio.com/")[[1]]
npkgs <- length(doe_pkgs)
```

```{r all_pkgs}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()%>% 
  mutate(Description = str_replace_all(Description, "\n", " "),
         Description = str_squish(Description),
         Title = str_replace_all(Title, "\n", " "))
```

```{r extra-pkgs}
#exclude_pkgs <- c("argonR", "card", "crispRdesignR", "genotypeR", "genepi", "recipes", "agridat", "desplot", "DiceView", "DOS", "DOS2", "EngrExpt")
cdoe_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | Package %in% doe_pkgs) %>% 
  #filter(!Title %in% exclude_pkgs) %>% 
  pull(Package) 
```

```{r}
# installation issues
exclude_pkgs <- c("osmgeosample", "secrdesign", "tanaka", "bioOED", "doebioresearch", "ssizeRNA", "SurvGSD")
design_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | 
           Package %in% doe_pkgs) %>% 
  mutate(taskview_doe = ifelse(Package %in% doe_pkgs, "yes", "no"),
         taskview_survey = ifelse(Package %in% survey_pkgs, "yes", "no"),
         criteria = case_when(str_detect(tolower(paste(Title, Description)),
                                         "(experiment|block design|factorial design|clinical trial|optimal design|response surface design|stepped wedge design|split-plot design|phase i|crossover design|cluster randomi[sz]ed|non-inferiority trial)") ~ "DoE",
                              TRUE ~ "No")) %>% 
  arrange(desc(criteria)) %>% 
  dplyr::select(Package, Title, Description, taskview_doe, taskview_survey, criteria) %>% 
  filter(!Package %in% exclude_pkgs)

write_csv(design_pkgs, file = "output/design_classification.csv")
```
```{r, include = FALSE, cache = FALSE}
pacman::p_load(char = design_pkgs$Package, install = FALSE, update = FALSE)
```

```{css, echo = FALSE}
.tooltip {
  position: relative;
  display: inline-block;
  border-bottom: 1px dotted black;
  opacity: 1;
}

.tooltip .tooltiptext {
  visibility: hidden;
  width: 120px;
  background-color: black;
  color: #fff;
  text-align: center;
  border-radius: 6px;
  padding: 5px 0;

  /* Position the tooltip */
  position: absolute;
  z-index: 1;
}

.tooltip:hover .tooltiptext {
  visibility: visible;
}
```

```{r}
# https://stackoverflow.com/questions/8918753/r-help-page-as-object
Rd2list <- function(Rd){
    names(Rd) <- substring(sapply(Rd, attr, "Rd_tag"),2)
    temp_args <- Rd$arguments

    Rd$arguments <- NULL
    myrd <- lapply(Rd, unlist)
    myrd <- lapply(myrd, paste, collapse="")

    temp_args <- temp_args[sapply(temp_args , attr, "Rd_tag") == "\\item"]
    temp_args <- lapply(temp_args, lapply, paste, collapse="")
    temp_args <- lapply(temp_args, "names<-", c("arg", "description"))
    myrd$arguments <- temp_args
    return(myrd)
}

getHelpList <- function(fn, pkg){
    thefile <- do.call("help", list(topic = fn, package = pkg))
    myrd <- utils:::.getHelpFile(thefile)
    Rd2list(myrd)$description %>% 
      str_replace_all("\\n", "") %>% 
      str_squish() %>% 
      str_trim() 
}
```

```{r}
design_pkgs %>% 
    filter(criteria == "DoE") %>% 
    mutate(fns = map(Package, ~as.character(lsf.str(glue("package:{.x}"))))) %>%
    mutate(func = map_chr(fns, ~paste(glue("<div class='tooltip'>{.x}<span class='tooltiptext'>getHelpList(x, {Package})</span></div>"), collapse = ", ")))   %>% 
  dplyr::select(-fns, -criteria) %>% 
  DT::datatable(escape = FALSE)

design_pkgs %>% 
  filter(criteria == "No") %>% 
  mutate(fns = map(Package, ~lsf.str(glue("package:{.x}")))) %>% 
  mutate(func = glue_collapse(map_chr(fns, function(x) glue("<div class='tooltip'>{x}<span class='tooltiptext'>{getHelpList(x, Package)}</span></div>")), ", ")) %>% 
  dplyr::select(-fns, -criteria) %>% 
  DT::datatable(escape = FALSE)

```

```{r}
fdoe_pkgs <- design_pkgs %>% 
  filter(criteria=="DoE") %>% 
  pull(Package)
```