### 2021-08-01 paper/rjournal/paper-rjournal.Rmd

---
author:
  # see ?rjournal_article for more information
  # Dai Dai to add her own name here
  - name: Emi Tanaka
    affiliation: Monash University
    address:
    - Monash University
    - Clayton campus, VIC 3800, Australia
    url: http://emitanaka.org/
    orcid: 0000-0002-1455-259X
    email:  emi.tanaka@monash.edu
preamble: |
  % Any extra LaTeX you need in the preamble

# per R journal requirement, the bib filename should be the same as the output 
# tex file. Don't forget to rename the bib file and change this example value.
bibliography: ../paper.bib

output: rticles::rjournal_article
---

```{r, include = FALSE}
paper <- here::here("paper/index.Rmd")
meta <- rmarkdown::yaml_front_matter(paper)
```

---
title: "`r meta$title`"
abstract: >
  `r meta$abstract`
---

```{r child=paper}
```