### 2021-08-16 analysis/doe-pkgs-analysis.Rmd
Transform the yaml into dataframe.

---
title: "doe-pkgs-analysis"
author: "Dewi Lestari Amaliah - 31251587"
date: "15/08/2021"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(yaml)
library(reshape)
library(tidyverse)
library(lubridate)
```

```{r}
#load the yml file and transform it as data frame
pkg_df <- yaml.load_file(here::here("data/input/doe-pkgs.yml")) %>%
  melt()

#tidy the data frame
pkg_tidy <- pkg_df %>%
  pivot_wider(names_from = L3,
              values_from = value)

# pivot longer the data frame by its type 
pkg_tidy_long_type <- pkg_tidy %>%
  unnest(type) %>%
  unnest(date) %>%
  unnest(domain) %>%
  mutate(year = year(ymd(date)))
```

```{r}
# summarise the packages based on its type
pkg_tidy_long_type %>%
  filter(relevance != "no") %>%
  group_by(type) %>%
  count() %>%
  arrange(desc(n))
```
```{r}
#summarise the package based on its date of published 
pkg_tidy_long_type %>%
  filter(relevance != "no") %>%
  group_by(year) %>%
  count() %>%
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  geom_point()
```

```{r}
pkg_tidy_long_type %>%
  filter(relevance != "no") %>%
  group_by(domain) %>%
  count() %>%
  arrange(desc(n))
```

```{r}
pkg_tidy_long_main <- pkg_tidy %>%
  unnest(main) %>%
  separate(main, into = c("funct_name", "funct_type", "funct_usage"), sep = ' - ')
```

```{r}
pkg_tidy_long_main %>%
  filter(relevance != "no") %>%
  group_by(funct_type) %>%
  count() %>%
  arrange(desc(n))
```