### 2021-09-12 paper/package_interconnection.Rmd
Add faceted network graph.

---
title: "Package Interconnection"
author: "Dewi Lestari Amaliah"
date: "03/09/2021"
output: bookdown::html_document2
bibliography: paper.bib
---

```{r setup-section4, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(yaml)
library(reshape)
library(tidyverse)
library(lubridate)
library(splitstackshape)
library(igraph)
library(tidygraph)
library(ggraph)
library(networkD3)
library(visNetwork)
library(qdapRegex)
```

# Interaction

```{r network-data}
pkg_df <- yaml.load_file(here::here("data/input/doe-pkgs.yml")) %>%
  melt()

#tidy the data frame
pkg_tidy <- pkg_df %>%
  pivot_wider(names_from = L3,
              values_from = value)

# make list of doe package 
list_pkg <- as.list(pkg_tidy)$L2

# create function to extract the depends, imports, and suggests 
get_pkg <- function(var){
  # select only the package name and depends/imports/suggests
  tmp_df <- pkg_tidy %>%
    select(L2, author, var) %>%
    # separate every depends/imports/suggests
    cSplit(var, sep = " ,")

  df <- tmp_df %>%
    pivot_longer(c(3:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
    filter(!is.na(pkg)) %>%
    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
    mutate(doe_pkg_related = ifelse(pkg %in% list_pkg, "yes", "no")) %>%
    # filter only package whose depends/imports/suggests is a DoE
    filter(doe_pkg_related == "yes") %>%
    # tidy the data to only have the type: depends/imports/suggests
    cSplit("type", sep = "_") %>%
    select(-type_2) %>%
    rename(type = type_1)

  return(df)
}

# create data frame from the above function
depend_df <- get_pkg("depends")
import_df <- get_pkg("imports")
suggest_df <- get_pkg("suggests")

# bind data frames
pkg_edges <- rbind(depend_df, import_df, suggest_df) %>%
  select(-doe_pkg_related) %>%
  select(pkg, L2, type, author)

# extract unique package that has depends/imports/suggests
from <- pkg_edges %>%
  distinct(L2)

# extract unique package that being depends/imports/suggests
to <- pkg_edges %>%
  distinct(pkg)

# create nodes from the two previous dataframe
pkg_nodes_tmp <- full_join(from, to, by = c("L2" = "pkg")) %>%
  left_join(pkg_tidy, by = "L2") %>%
  select(L2, author)

### get author and count how many packages they write as the first author
### there are many authors who wrote >1 packages but in the same packages, here, we just extract the first author

# get first author 
pkg_nodes_tmp$author <- rm_angle(pkg_nodes_tmp$author, replacement = "")
pkg_nodes_tmp$author <- rm_square(pkg_nodes_tmp$author, replacement = "")
pkg_nodes_tmp$author <- rm_white_punctuation(pkg_nodes_tmp$author, replacement = "")
pkg_nodes_tmp$author <- ifelse(str_detect(pkg_nodes_tmp$author, "based"), "Ernesto Barrios", pkg_nodes_tmp$author)

pkg_nodes_tmp <- pkg_nodes_tmp %>%
  cSplit("author", sep = " ,") %>%
  select(c(1:2)) %>%
  rename(first_aut = author_1)

aut_gr <- pkg_nodes_tmp %>%
  group_by(first_aut) %>%
  count() %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  right_join(pkg_nodes_tmp, by = "first_aut")

pkg_nodes <- aut_gr %>%
  mutate(first_author = case_when(first_aut == "Ulrike Groemping" ~ "Ulrike Groemping",
                                 first_aut == "John Lawson" ~ "John Lawson",
                                 first_aut == "Robert B. Gramacy" ~ "Robert B. Gramacy",
                                 first_aut == "Ernesto Barrios" ~ "Ernesto Barrios",
                                 first_aut == "Kevin Wright" ~ "Kevin Wright",
                                 first_aut == "Shwetank Lall" ~ "Shwetank Lall",
                                 n == 1 ~ "Other")) %>%
  mutate(first_author = factor(first_author, levels = c("Ulrike Groemping",
                                                        "John Lawson",
                                                        "Robert B. Gramacy",
                                                        "Ernesto Barrios",
                                                        "Kevin Wright",
                                                        "Shwetank Lall",
                                                        "Other"))) %>%
  select(L2, first_author)

# create igraph object
routes_igraph <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes, directed = TRUE)

# create the input based on the igraph object
routes_tidy <- routes_igraph %>% as_tbl_graph() %>%
  activate(edges)

# create the nodes for interactive network
nodes <- pkg_nodes %>%
  rowid_to_column("id") %>%
  rename(label = L2) %>%
  mutate(font.size = 25)

#create edges for interactive network
edges <-routes_tidy %>%
  activate(edges) %>%
  data.frame()

edges$color <- case_when(edges$type == "imports" ~ "green",
                         edges$type == "depends" ~ "red",
                         edges$type == "suggests" ~ "blue")

ledges <- data.frame(color = c("green", "red", "blue"), 
                     label = c("imports", "depends", "suggests"))

```

Interaction between packages could indicate of how developers utilize the functionality of one or more packages to create new features in their packages. According to @StanstrupJan2019TmTi, dependency and connectivity in terms of reusing codes and inheriting objects could reflect a well-connected and interacting community. In this paper, we observe the interaction limited to only between packages that are listed as DoE packages. Further, the type of interaction include the depended, imported, and suggested packages. Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes)` packages that interplay each other. To observe these interaction, we use network visualisation using `igraph` [@igraph], `tidygraph` [@tidygraph], and `ggraph` [@ggraph].

```{r network-graph, fig.cap = "Interaction network between DoE R packages. The color of edges reflect the type of interaction, i.e., depends, imports, and suggests. Nodes' color indicates the authors who wrote more than 2 DoE packages as the first author.", fig.width=12, fig.height = 8}

# create the graph
ggraph(routes_tidy, layout = "kk") + 
  geom_edge_link(aes(color = type),
                 arrow = arrow(length = unit(2, 'mm')), 
                 end_cap = circle(2, 'mm')) + 
  geom_node_point(aes(color = first_author), size = 4) + 
  geom_node_text(aes(label = name), repel = TRUE, size = 3.1) +
  scale_color_manual(values = c("blue", "tomato", "magenta", "grey50", "#48C9B0", "#C0392B", "#F1C40F")) +
  theme_graph() 
```

```{r, fig.width=12, fig.height = 6}

pkg_nodes_facet <- pkg_nodes %>%
  filter(first_author %in% c("Ulrike Groemping", "John Lawson", "Robert B. Gramacy"))

pkg_edges_facet <- pkg_edges %>%
  filter(L2 %in% pkg_nodes_facet$L2 &
           pkg %in% pkg_nodes_facet$L2)

routes_igraph_facet <- graph_from_data_frame(d = pkg_edges_facet, 
                                             vertices = pkg_nodes_facet, 
                                             directed = TRUE)

# # create the input based on the igraph object
routes_tidy_facet <- routes_igraph_facet %>% as_tbl_graph() %>%
  activate(edges)

ggraph(routes_tidy_facet, layout = "kk") + 
  geom_edge_link(aes(color = type),
                 arrow = arrow(length = unit(3, 'mm')), 
                 end_cap = circle(3, 'mm')) + 
  geom_node_point(aes(color = first_author), size = 4) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3.5) +
  theme_graph() +
  facet_nodes(~first_author) +
  scale_color_manual(values = c("tomato", "#48C9B0", "#F1C40F"), guide = "none") +
  th_foreground(foreground = "grey80",  border = TRUE) 
```

```{r int-plot}
visNetwork(nodes, edges, height = "500px", width = "100%") %>%
  visOptions(selectedBy = "first_author") %>%
  visLayout(randomSeed = 123) %>%
  visEdges(shadow = FALSE,
           arrows =list(to = list(enabled = TRUE, scaleFactor = 2))) %>%
  visLegend(useGroups = F, addEdges = ledges)
```

Figure \@fig(network-graph) shows the interplay of DoE packages. In the previous section, we noticed that Ulrike Groemping authored many DoE packages. We also noticed that one of her package, DoE base, is among the most downloaded DoE package on CRAN. Hence, we categorise the package whether it is authored by her or not. The network conveys that there are 5 clusters of interaction with one big cluster.

The most downloaded DoE packages, such as `AlgDesign`, `rsm`, and `lhs` are observed to support many other packages. Dependent packages seems to use their functions to support their feature. We also observe that  `DoE.base` and `FrF2`, which are authored by Groemping, are also imported, depended, and suggested by many other packages because of their basic functionality in experimental design. Moreover, some of interacting packages, such as `DoE.wrapper` and `DoE.MIParray` are also developed by Groemping.

Further, packages regarding computer experiment (`DiceKriging`, `DiceEval`, `DiceView`, and `DiceDesign`) happen to interplay with each other so that they appear to form a sub-cluster. The four smaller clusters interact because of the similarity in their type and domain. For example `dfcm` and `dfpk` are used for dose-finding experiments. Regarding the similarity of its objective, although `FMC` aims to generate factorial design, this package does not interact with the packages in the big cluster, even though some packages in this cluster, such as `DoE.base`, also involve factorial design. This package instead imports `minimalRSD`, which is developed by the same author.

Based on this network, we can say that the interconnection of DoE package is due to functionality dependence, similarity of type or domain, and because they are developed by the same author/s.