### 2021-10-24 tidy_author.Rmd
Add author tidying file.

---
title: "tidy_author"
author: "Dewi Amaliah"
date: "24/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(qdapRegex)
library(entity)
```

```{r}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()

ctvs <- map_dfr(ctv::available.views(repos = "http://cran.rstudio.com/"), 
                  ~tibble(name = .x[["name"]],
                          topic = .x[["topic"]],
                          pkgs = map(name, ~ctv:::.get_pkgs_from_ctv_or_repos(.x, 
                                                repos = "http://cran.rstudio.com/")[[1]])))

ctvdf <- ctvs %>% 
  unnest_longer(pkgs) %>% 
  left_join(select(db, Package, Author, Depends, Imports, Suggests), 
            by = c("pkgs" = "Package")) %>% 
  mutate(across(c(Depends, Imports, Suggests),
                ~str_split(.x, ", "))) %>% 
  mutate(across(c(Depends, Imports, Suggests),
                ~map(.x, function(apkg) str_extract(apkg, "^[A-Za-z0-9]+")))) %>%
  # remove angle bracket
  mutate(author_tidy = rm_angle(Author, replacement = ""),
         # remove square bracket
         author_tidy = rm_square(author_tidy, replacement = ""),
         # remove bracket
         author_tidy = rm_bracket(author_tidy, replacement = ""),
         author_tidy = str_replace(author_tidy, "&", ","),
         author_tidy = str_replace(author_tidy, '\\\\', ""),
         # remove white space before comma
         author_tidy = rm_white_comma(author_tidy, replacement = ""),
         # replace "and" with comma
         author_tidy = str_replace(author_tidy, " and", ","),
         # remove extra comma 
         author_tidy = str_replace(author_tidy, ",,", ","),
         author_tidy2 = ifelse(str_detect(author_tidy, "riginal") | 
                                 str_detect(author_tidy, " by") | 
                                 str_detect(author_tidy, "with") | 
                                 str_detect(author_tidy, " under"), 
                               person_entity(author_tidy), 
                               author_tidy))

ctvdf_sentences <- ctvdf %>%
  filter(str_detect(author_tidy, "riginal") | str_detect(author_tidy, " by") | str_detect(author_tidy, "with") | str_detect(author_tidy, " under"))

ctvdf_sentences_test <- ctvdf_sentences %>%
  mutate(author_tidy2 = person_entity(author_tidy)) %>%
  mutate(author_tidy3 = str_replace(author_tidy, "with contributions from", ","),
         author_tidy3 = str_replace(author_tidy, "Original by", ""))

ctvdf_with <- ctvdf %>%
  filter(str_detect(author_tidy, "with"))

# try to filter English words

print(ctvdf_with$author_tidy)
```