### 2022-05-31 other/tidy_author-ET.Rmd
significant update to the paper

---
title: "tidy_author"
author: "Dewi Amaliah"
date: "24/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(qdapRegex)
#library(entity)
```

```{r}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()

ctvs <- map_dfr(ctv::available.views(repos = "http://cran.rstudio.com/"), 
                  ~tibble(name = .x[["name"]],
                          topic = .x[["topic"]],
                          pkgs = map(name, ~ctv:::.get_pkgs_from_ctv_or_repos(.x, 
                                                repos = "http://cran.rstudio.com/")[[1]])))

words_to_delete <- c(" with", " by", " under", " contributions?", " is", " the", 
                     " author", " [Oo]riginal(ly)?", " C")

ctvdf <- ctvs %>% 
  unnest_longer(pkgs) %>% 
  left_join(select(db, Package, Author, Depends, Imports, Suggests), 
            by = c("pkgs" = "Package")) %>% 
  mutate(across(c(Depends, Imports, Suggests),
                ~str_split(.x, ", "))) %>% 
  mutate(across(c(Depends, Imports, Suggests),
                ~map(.x, function(apkg) str_extract(apkg, "^[A-Za-z0-9]+")))) %>%

  mutate(
    # remove angle bracket
    author_tidy = str_replace_all(Author, "\\n", ""),
    # remove square bracket
    author_tidy = str_replace_all(author_tidy, "<.+>", ""),
    author_tidy = str_replace_all(author_tidy, "\\[.+\\]", ""),
    # remove bracket
    author_tidy = str_replace_all(author_tidy, "\\(.*\\)", ""),
    author_tidy = str_replace_all(author_tidy, "[.]", ""),
    author_tidy = str_replace_all(author_tidy, "(&| and)", ","),
    author_tidy = str_replace_all(author_tidy, ",,", ","),
    author_tidy = str_replace(author_tidy, " ,$", ""),
    author_tidy = str_replace_all(author_tidy, 
                                  paste0("(", paste(words_to_delete, collapse = "|"), ")"),
                                  ""),
    author_tidy = str_squish(author_tidy)) %>% 
  select(Author, author_tidy)

tidytext::stop_words

ctvdf_sentences <- ctvdf %>%
  filter(str_detect(author_tidy, "riginal") | str_detect(author_tidy, " by") | str_detect(author_tidy, "with") | str_detect(author_tidy, " under"))

ctvdf_sentences_test <- ctvdf_sentences %>%
  mutate(author_tidy2 = person_entity(author_tidy)) %>%
  mutate(author_tidy3 = str_replace(author_tidy, "with contributions from", ","),
         author_tidy3 = str_replace(author_tidy, "Original by", ""))

ctvdf_with <- ctvdf %>%
  filter(str_detect(author_tidy, "with"))

# try to filter English words

print(ctvdf_with$author_tidy)
```