### 2022-06-18 README.Rmd
make it easier for people to navigate the repo

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)
library(tidyverse)
paper <- here::here("paper.Rmd")
meta <- rmarkdown::yaml_front_matter(paper)
```

# `r meta$title`

This is the repo for the paper titled "`r meta$title`".

📜 You can find the arxiv version of the paper [here](https://arxiv.org/abs/2206.07532).

The paper uses the developmental version of the [`cranscrub` package in the branch `paper-doe-review`](https://github.com/numbats/cranscrub/tree/paper-doe-review). This can be installed by

```r
remotes::install_github("numbats/cranscrub", ref = "paper-doe-review")
```

To reproduce the results of the paper, you will need to run the analysis by calling

```r
targets::tar_make()
```

The primary functions for analysis are:

* `R/functions.R`
* `R/plot.R`
* `_targets.R`

The contents of the paper can be found in `paper.Rmd` and the supplementary materials in `supp.Rmd`.

## Authors

```{r, echo = FALSE}
map_chr(meta$author, ~{
  paste0("- ", .x$name, ", ", .x$affiliation)
}) %>% 
  cli::cli_li() 
```

## Abstract

`r meta$abstract`

## Supplementary material

The supplementary material can be found [here](https://emitanaka.org/paper-DoE-review/supp.pdf).