### 2022-06-18 other/analysis/analysis-download-logs.Rmd

---
title: "Download logs"
author: "Emi Tanaka"
date: "24/03/2021"
output: html_document
---

```{r library, include=FALSE}
library(tidyverse)
library(magrittr)
library(rvest)
library(glue)
library(lubridate)
library(plotly)
library(cranlogs)
library(ggbeeswarm)
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache.path = "cache/")
```

```{r cache, eval = FALSE}
fs::dir_ls("cache") %>% 
  str_subset(".RData$") %>% 
  walk(load)
```

```{r ctv}
doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
                                              repos = "http://cran.rstudio.com/")[[1]]
survey_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
                                              repos = "http://cran.rstudio.com/")[[1]]
npkgs <- length(doe_pkgs)
```

```{r all_pkgs}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()%>% 
  mutate(Description = str_replace_all(Description, "\n", " "),
         Description = str_squish(Description),
         Title = str_replace_all(Title, "\n", " "))
```

```{r extra-pkgs}
exclude_pkgs <- c("argonR", "card", "crispRdesignR", "genotypeR", "genepi", "recipes",
                  "agridat", "desplot", "DiceView", "DOS", "DOS2", "EngrExpt")
cdoe_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | Package %in% doe_pkgs) %>% 
  filter(!Title %in% exclude_pkgs) %>% 
  pull(Package) 
```

```{r}
fdoe_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | 
           Package %in% doe_pkgs) %>% 
  mutate(taskview_doe = ifelse(Package %in% doe_pkgs, "yes", "no"),
         taskview_survey = ifelse(Package %in% survey_pkgs, "yes", "no"),
         criteria = case_when(str_detect(tolower(paste(Title, Description)),
                                         "(experiment|block design|factorial design|clinical trial|optimal design|response surface design|stepped wedge design|split-plot design|phase i|crossover design|cluster randomi[sz]ed|non-inferiority trial)") ~ "DoE",
                              TRUE ~ "No")) %>% 
  mutate(construct = case_when(str_detect(tolower(paste(Title, Description)), "construct|comput(e|ing)|creat(e|ing)|find(ing)") ~ "yes",
                               TRUE ~ "no"
                                          )) %>% 
  arrange(desc(criteria), construct) %>% 
  select(Package, Title, Description, taskview_doe, taskview_survey, criteria, construct) %T>% 
  write_csv(file = "output/design_classification.csv") %>% 
  filter(criteria=="DoE") %>% 
  pull(Package)
```

```{r}
Rd2list <- function(Rd){
    names(Rd) <- substring(sapply(Rd, attr, "Rd_tag"),2);
    temp_args <- Rd$arguments;

    Rd$arguments <- NULL;
    myrd <- lapply(Rd, unlist);
    myrd <- lapply(myrd, paste, collapse="");

    temp_args <- temp_args[sapply(temp_args , attr, "Rd_tag") == "\\item"];
    temp_args <- lapply(temp_args, lapply, paste, collapse="");
    temp_args <- lapply(temp_args, "names<-", c("arg", "description"));
    myrd$arguments <- temp_args;
    return(myrd);
}

getHelpList <- function(fn, pkg){
    thefile <- do.call("help", list(topic = fn, package = pkg))
    myrd <- utils:::.getHelpFile(thefile);
    Rd2list(myrd)$description %>% 
      str_replace_all("\\n", "") %>% 
      str_squish() %>% 
      str_trim() 
}
pkg <- "agricolae"
library(pkg, character.only = TRUE)
fn <- lsf.str(glue::glue("package:{pkg}"))[2]
getHelpList(fn, pkg)
```

```{r cran-data}
end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
start <- end - years(2) + days(2)
dldat <- cran_downloads(fdoe_pkgs, from = start, to = end)
```

```{r summary}
dldat_sum <- dldat %>% 
  group_by(package) %>% 
  summarise(total = sum(count)) %>% 
  arrange(desc(total)) %>% 
  mutate(rank = 1:n(),
         taskview = package %in% doe_pkgs)
```

```{r graph}
g1 <- ggplot(dldat_sum, aes(x = taskview, y = total/1000, label = package, text = rank, color = taskview)) + 
  geom_beeswarm() +
  scale_y_log10(label = comma) + 
  labs(x = "",
       y = "Download counts in the last 2 year ('000s)") + 
  theme(axis.ticks.length.x = unit(0, "mm"))

ggplotly(g1)
```

```{r updates}
pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
pkg_updates <- map(fdoe_pkgs, function(pkg) {
    last_update <- read_html(glue(pkg_url)) %>% 
      html_table() %>% 
      .[[1]] %>% 
      filter(X1=="Published:") %>% 
      pull(X2) %>% 
      ymd()

    archive_dates <- tryCatch({ 
        read_html(glue(pkg_archive)) %>% 
          html_table() %>%
          .[[1]] %>% 
          pull(`Last modified`) %>% 
          ymd_hm() %>% 
          na.omit() %>% 
          as.Date()
      }, error = function(e) {
        NULL
      })
    c(archive_dates, last_update)
  })
names(pkg_updates) <- doe_pkgs

updates <- unlist(pkg_updates) %>% 
  enframe("package", "update") %>% 
  # unlist converts date to integers
  mutate(update = as.Date(update, origin = "1970-01-01"),
         # need to get rid of the numbers appended to pkg names
         package = str_extract(package, paste0(doe_pkgs, collapse="|"))) 
```

```{r first}
pkgsum <- updates %>% 
  group_by(package) %>% 
  summarise(first = min(update),
            nupdates = n()) %>% 
  left_join(dldat_sum, by = "package")

g2 <- ggplot(pkgsum, aes(first, total, label = package, text = rank, size = nupdates)) + 
  geom_point(alpha = 0.4) + 
  scale_y_log10() + 
  #geom_smooth() +
  NULL
ggplotly(g2)
```

* Packages released after 2017 September doesn't seem to be listed in CRAN Task View

```{r}
g3 <- ggplot(pkgsum, aes(nupdates, total, label = package, text = rank)) + 
  geom_point(alpha = 0.4) + 
  scale_y_log10() + 
  scale_x_log10() + 
  #geom_smooth() +
  NULL
ggplotly(g3)
```